---
title: "Lab 7: Simple Functions and Conditionals"
subtitle: "Stat 133, Spring 2018"
author: "Kanak Garg"
output: html_document
fontsize: 12pt
urlcolor: blue
---

>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```


By setting the global option `error = TRUE` you avoid the knitting process 
to be stopped in case a code chunk generates an error.

Since you willl be writing a couple of functions with `stop()` statements,
it is essential that you set up `error = TRUE`, otherwise `"knitr"` will 
stop knitting your `Rmd` file if it encounters an error.

## Simple Math Functions

Consider the following mathematical functions:

- $f(x) = x^2$
- $g(x) = 2x + 5$

Write two functions `f()` and `g()` based on the previous equations. Don't
forget to include roxygen comments to document your function!

```{r}

# @title square numbers
# @description square numbers
# @param int x
# @return square of x

f <- function(x = 0) {
  return(x*x)
}

```

```{r}
# @title function
# @description apply g on numbers
# @param int x
# @return g(x)

g <- function(x = 0) {
  return(2*x + 5)
}

```


Test your functions with:

```{r}
f(2)     # 4
f(-5)    # 25

g(0)     # 5
g(-5/2)  # 0
```

Use your functions `f()` and `g()` to create the following composite functions:

- `fog()` for the composite function: $f \circ g(x)$
- `gof()` for the composite function: $g \circ f(x)$

```{r}
# @title f of g of x
# @description f of g of x
# @param int x
# @return f of g of x

fog <- function(x = 0) {
  return(f(g(x)))
}


```

```{r}
# @title g of f of x
# @description g of f of x
# @param int x
# @return g of f of x

gof <- function(x = 0) {
  return(g(f(x)))
}

```

Test your composite functions with:

```{r}
fog(2)     # 81
fog(-5)    # 25

gof(0)     # 5
gof(-5/2)  # 17.5
```


-----


## Pythagoras

The pythagoras formula is used to compute the length of the hypotenuse, $c$, of 
a right triangle with legs of length $a$ and $b$.

![hypotenuse](https://wikimedia.org/api/rest_v1/media/math/render/svg/5fd521cee81d583ce94bf6710984cc2a9eb7c3da)

Write a function `pythagoras()` that takes two arguments `a` and `b`, and 
returns the length of the hypotenuse. Don't forget to include roxygen comments 
to document your function!

```{r}
# @title pythagoraas
# @description find c
# @param int a,b
# @return c

pythagoras <- function(a = 1, b =a) {
  return(sqrt(f(a)+f(b)))
}
```

Test your `pythagoras()` with two leg values: `pythagoras(3, 4)`

```{r}
pythagoras(3, 4)

```

Modify your function `pythagoras()` so that argument `b` takes the same value of argument `a`. Test it with just one leg value: `pythagoras(5)`

```{r}
pythagoras(5)

```


-----

## Area of a circle

Consider a circle with `radius = 2`. The area of this circle can be computed
in R as:

```{r}
# @title area of circle
# @description area of circle
# @param int r
# @return area

circle_area <- function(radius = 1) {
  return(pi*f(radius))
}
```

Write a function `circle_area()` that calculates the area of a circle. This 
function must take one argument `radius`. Give `radius` a default value of 1.
Don't forget to include roxygen comments to document your function!

For example:

```{r eval = FALSE}
# default (radius 1)
circle_area()

# radius 3
circle_area(radius = 3)
```

Modify your `circle_area()` function in order to include a `stop()` statement.
If `radius` is negative, then the function should stop with a message like:
`"radius cannot be negative"`.

Test your modified `circle_area()` with `radius = -2`; the function should 
return a stop message:

```{r, eval = FALSE}
# bad radius
circle_area(radius = -2)
```



## Area of a cylinder

For a given cylinder of radius $r$ and height $h$ the area $A$ is:

$$
A = 2 \pi r h + 2 \pi r^2
$$

For example. Say you have a cylinder with radius = 2, and height = 3.

```{r}

# @title area
# @description area
# @param int r,h
# @return area

cylinder_area <- function(radius = 1, height=1) {
  if(radius<0 | height <0){
    return("error")
  }
  return(2 * pi * radius * height + 2* circle_area(radius))
}

```


Notice that the formula of the area of a cylinder includes the area of a 
circle: $\pi r^2$. Write a function `cylinder_area()`, that calls `circle_area()`, 
to compute the area of a cylinder. 

This function must take two arguments: `radius` and `height`. Give both 
arguments a default value of 1. In addition, the function should stop if any
of `radius` or `height` are negative.

For instance:

```{r eval = FALSE}
# default (radius 1, height 1)
cylinder_area()

# radius 2, height 3
cylinder_area(radius = 2, height = 3)
```

These should return an error message:

```{r, eval = FALSE}
# bad radius
cylinder_area(radius = -2, height = 1)

# bad height
cylinder_area(radius = 2, height = -1)

# bad radius and height
cylinder_area(radius = -2, height = -1)
```


## Volume of a cylinder

For a given cylinder of radius $r$ and height $h$ the volume $V$ is:

$$
V = \pi r^2 h
$$

Write a function `cylinder_volume()`, that calls `circle_area()`, to compute the 
volume of a cylinder. This function must take two arguments: 
`radius` and `height`. Give both arguments a default value of 1.
```{r}

# @title volume
# @description volume
# @param int r,h
# @return volume

cylinder_volume <- function(radius = 1, height=1) {
  if(radius<0 | height <0){
    return("error")
  }
  return( 2* circle_area(radius) * height )
}

```

For example:

```{r eval = FALSE}
# default (radius 1, height 1)
cylinder_volume()

cylinder_volume(radius = 3, height = 10)
cylinder_volume(height = 10, radius = 3)
```



### Miles to Kilometers

Write a function `miles2kms()` that converts miles into kilometers: 1 mile
is equal to 1.6 kilometers. Give the argument a default value of 1.

```{r}
miles2kms <- function( m = 1) {
  return(1.6*m)
}
```




### Polynomials

In this problem we want to see whether the graph of a given polynomial
will cross or touch the _x_ axis in a given interval. 

Let's begin with the polynomial: $f(x) = x^2 (x - 1)$. The first thing 
to do is write a function for the polynomial, for instance:

```{r}
poly1 <- function(x) {
  (x^2) * (x - 1)
}
```

Once you have a function for the polynomial, you can create a set of pairs of
points $x$ and $y = f(x)$, and then use them to graph the polynomial

```{r polynomial_plot}
# set of points
x <- seq(-4, 4, length.out = 20)
y <- poly1(x)

# graph polynomial
plot(x, y, type = 'l', lwd = 3, col = "#FB7215", las = 1)
abline(h = 0, v = 0, col = '#888888aa', lwd = 1.5)
title(main = expression(paste(f(x), ' = ', x^2, (x - 1))))
```

```{r}
poly1 <- function(x) {
  (x^2 - 1)*(x + 3)^3
}
```

```{r}
# set of points
x <- seq(-4, 4, length.out = 20)
y <- poly1(x)

# graph polynomial
plot(x, y, type = 'l', lwd = 3, col = "#FB7215", las = 1)
abline(h = 0, v = 0, col = '#888888aa', lwd = 1.5)
title(main = expression(paste(f(x))))
```

-----


## Descriptive Statistics

Write a function `descriptive()` that takes a numeric vector as input, and 
returns a named vector with the following descriptive statistics:

- `min`: minimum
- `q1`: first quartile (Q2)
- `median`: median
- `mean`: mean
- `q3`: third quartile (Q3)
- `max`: maximum
- `range`: range or span (max - min)
- `iqr`: interquartile range (IQR)
- `sd`: standard deviation

```{r}
# your descriptive() function
descriptive <- function( vect) {
  return(c(min(vect), median(vect), mean(vect), max(vect), range(vect), IQR(vect), sd(vect)))
}
```

