---
title: "Lab 4: Data Frame Basics"
subtitle: "Stat 133, Spring 2018"
author: "Kanak Garg"
output: github_document
urlcolor: blue
---
## Abalone Data Set

```{r}
url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
abalone <- read.table(url, sep = ",")
```

## Basic Importing

```{r}
abalone <- read.table("abalone.data", sep = ",")
```


```{r}
head(abalone)

tail(abalone)
```

```{r}
str(abalone, vec.len = 1)
```


### Detailed information about the columns 


```{r}

column_names <- c(
	'sex',
	'length',
	'diameter',
	'height',
	'whole_weight',
	'shucked_weight',
	'viscera_weight',
	'shell_weight',
	'rings'
)

column_types <- c(
	'character',
	'real',
	'real',
	'real',
	'real',
	'real',
	'real',
	'real',
	'integer'	
)
```


```{r}
abalone <- read.table(
	'abalone.data',
	col.names = column_names,
	colClasses = column_types,
	sep = ","
)


str(abalone, vec.len = 1)
```


### Your turn

- Read the Abalone data with the `read.csv()` function.
- Use the inputs `col.names` and `colClasses` to specify column names and their data types.
- Look at the data description in the following link:
```{r}
read.csv("abalone.data")
```
```{r}
summary(abalone)
```

## NBA Players Data

```{r}
csv <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/data/nba2017-players.csv"
download.file(url = csv, destfile = 'nba2017-players.csv')
dat <- read.csv('../data/nba2017-players.csv', stringsAsFactors = FALSE)
```

-----


### Your turn:

Use bracket notation, the dollar operator, as well as concepts of 
logical subsetting and indexing to:

- Display the last 5 rows of the data.

```{r}
tail(dat)
```


- Display those rows associated to players having height less than 70 inches tall.

```{r}
dat1 <- dat[dat$height < 70, ]
dat1
```


- Of those players that arloe centers (position `C`), display their names and salaries.

```{r}
dat2 <- dat[dat$position == 'C', c(1,9) ]
dat2
```

- Create a data frame `durant` with Kevin Durant's information (i.e. row).
```{r}
durant <- data.frame(list(dat[dat$player == 'Kevin Durant', ]))
durant
```

- Create a data frame `ucla` with the data of players from college UCLA ("University of California, Los Angeles").
```{r}
ucla <- data.frame(list(dat[dat$college == 'University of California, Los Angeles', ]))
ucla
```

- Create a data frame `rookies` with those players with 0 years of experience. 
```{r}
rookies <- data.frame(list(dat[dat$experience == '0', ]))
rookies
```

- Create a data frame `rookie_centers` with the data of Center rookie players. 
```{r}
rookie_centers <- data.frame(list(dat[dat$experience == '0' & dat$position == "C", ]))
rookie_centers
```


- Create a data frame `top_players` for players with more than 50 games and more than 100 minutes played.

```{r}
top_players <- data.frame(list(dat[dat$games > 50 & dat$minutes > 100, ]))
top_players
```


- What's the largest height value?
```{r}
max(dat$height)
```

- What's the minimum height value?
```{r}
min(dat$height)
```


- What's the overall average height?
```{r}
mean(dat$height)
```


- Who is the tallest player?
```{r}
dat[dat$height == max(dat$height), ]
```


- Who is the shortest player?
```{r}
dat[dat$height == min(dat$height), ]
```

- Which are the unique teams?
```{r}
unique(dat$team)
```

- How many different teams?
```{r}
length(unique(dat$team))
```

- Who is the oldest player?
```{r}
dat[dat$age == max(dat$age), ]
```

- What is the median salary of all players?
```{r}
dat[dat$salary == median(dat$salary), ]
```

- What is the median salary of the players with 10 years of experience or more?
```{r}
dat3 <- dat[dat$experience >= 10, ]
median(dat3$salary)
```

- What is the median salary of Shooting Guards (SG) and Point Guards (PG)?
```{r}
dat4 <- dat[dat$position == "SG" | dat$position == "PG", ]
median(dat4$salary)
```

- What is the median salary of Power Forwards (PF), 29 years or older, and 74 inches tall or less?
```{r}
dat5 <- dat[dat$position == "PF" & dat$age >= 29  & dat$height <= 74, ]
median(dat5$salary)
```

- How many players scored 4 points or less?
```{r}
dat6 <- dat[dat$points <= 4, ]
length(dat6$player)
```

- Who are those players who scored 4 points or less?
```{r}
dat6$player
```


- Who is the player with 0 points?
```{r}
dat7 <- dat[dat$points == 0, ]
dat7$player
```

- How many players are from "University of California, Berkeley"?
```{r}
dat8 <- dat[dat$college == 'University of California, Berkeley', ]
length(dat8$player)
```


- Are there any players from "University of Notre Dame"? If so how many and who are they?
```{r}
dat8 <- dat[dat$college == 'University of Notre Dame', ]
dat8
```

- Are there any players with weight greater than 260 pounds? If so how many and who are they?
```{r}
dat8 <- dat[dat$weight >= 260, ]
dat8
```

- Create a data frame `gsw` with the name, height, weight of Golden State Warriors (GSW)

```{r}
gsw <- dat[dat$team == "GSW", c(1,4,5) ]
gsw
```


- Display the data in `gsw` sorted by height in increasing order (hint: see `?sort` and `?order`)
```{r}
gsw[order(gsw$height), ]
```

- Display the data in gsw by weight in decreasing order (hint: see `?sort` and `?order`)
```{r}
gsw[order(gsw$weight, decreasing = TRUE), ]
```

### Your turn

- Create a data frame with the average height, average weight, and average 
age, grouped by position

```{r}
aggregate(dat[, c("height", "weight", "age")], by = list(dat$position), FUN = mean)
```
- Create a data frame with the average height, average weight, and average 
age, grouped by team

```{r}
aggregate(dat[, c("height", "weight", "age")], by = list(dat$team), FUN = mean)
```
- Create a data frame with the average height, average weight, and average 
age, grouped by team and position.

```{r}

aggregate(dat[, c("height", "weight", "age")], by = list(dat$position, dat$team), FUN = mean)

```

