---
title: "Lab 2: Vectors and other data structures"
subtitle: "Stat 133, Spring 2018"
author: "Kanak Garg"
---
### Getting the Data File
```{r}
rm(list = ls())
rdata <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/data/nba2017-salary-points.RData"
download.file(url = rdata, destfile = 'nba2017-salary-points.RData')
load('nba2017-salary-points.RData')
```
```{r}
ls()
```

All the data values have the same length with 441 values None of the objects contain any null/NA values in their list. All the data objects except for "team," are all vectors. Their data types can be determined from above, whether they are a logical vector, character vector or a numeric vector. Team is a factor. 

## Manipulating Vectors: Subsetting

- all the even elements in `player`
```{r}
player[seq(2, length(player), 2)]
```

- all the odd elements in `salary`
```{r}
salary[seq(1, length(salary), 2)]
```

- all multiples of 5 (e.g. 5, 10, 15, etc) of `team`
```{r}
team[seq(5, length(team), 5)]
```

- elements in positions 10, 20, 30, 40, etc of `scored`
```{r}
points[seq(10, length(points), 10)]
```

- all the even elements in `team` but this time in reverse order
```{r}
rev(team[seq(2, length(team), 2)])
```

- players in position Center, of Warriors (GSW)
```{r}
player[team == "GSW" & position == "C"]
```

- players of both GSW (warriors) and LAL (lakers)
```{r}
player[team == "GSW" & team == "LAL"]
```

- players in positions Shooting Guard and Point Guards, of Lakers (LAL) 
```{r}
player[team == "LAL" & (position == "SG" | position == "PG")]

```

- subset Small Forwards of GSW and LAL
```{r}
player[team == "LAL" | team == "GSW" & position == "SF"]
```

- name of the player with largest salary
```{r}
player[salary == max(salary)]
```

- name of the player with smallest salary
```{r}
player[salary == min(salary)]
```

- name of the player with largest number of scored points
```{r}
player[points == max(points)]
```

- salary of the player with largest number of points
```{r}
salary[points == max(points)]
```

- largest salary of all Centers
```{r}
max(salary[position == "C"])
```

- team of the player with the largest number of scored points
```{r}
team[player == player[points == max(points)]]
```

- name of the player with the largest number of 3-pointers
```{r}
player[points3 == max(points3)]
```



## Subsetting with Character Vectors

```{r}
warriors_players <-  player[team == "GSW"]
warriors_salary <- salary[team == "GSW"]
warriors_points <- points[team == "GSW"]
names(warriors_salary) <-  warriors_players
```

### Some plotting

```{r}
plot(warriors_points, warriors_salary)
text(warriors_points, warriors_salary, labels = abbreviate(player))
```

```{r}
log_wp <- log(warriors_points)
log_ws <- log(warriors_salary)
plot(log_wp, log_ws)
text(log_wp, log_ws, labels = abbreviate(player))
```



## Recycling


`salary` is given in dollars, but what if you need to
obtain the salaries in millions of dollars?. Create a new vector
`salary_millions` with the converted values in millions of dollars. 

```{r}
salary_millions <- salary/1000000
```

Take the values in `points1`, `points2`, and `points3` and figure out to create  a new vector `scored_points` that gives you the same values of `points`.
```{r}
scored_points = points1 + 2 * points2 + 3* points3
```

### Creating Factors

```{r}
position_fac <- factor(position)

table(position_fac)
```


### Manipulating Factors
- positions of Warriors
```{r}
position_fac[team == "GSW"]
```

- positions of players with salaries > 15 millions
```{r}
position_fac[salary > 15000000]
```

- frequencies (counts) of positions with salaries > 15 millions
```{r}
table(position_fac[salary > 15000000])
```

- relative frequencies (proportions) of 'SG' (Shooting Guards) in each team
```{r}
table(team[position == "SG"])/table(team[team == team])

```


### More Plots
```{r}
plot(warriors_points, warriors_salary, col = position_fac )
text(warriors_points, warriors_salary)

```
