---
title: "Lab 9: Testing functions, and string basics"
subtitle: "Stat 133, Spring 2018"
author: "Kanak Garg"
---


```{r}
library(testthat)
```


# Basics of String Manipulation

In this second part of the lab,  you will be using the row names of the data frame `USArrests`
(this data comes already in R):

```{r usarrests}
head(USArrests)

states <- rownames(USArrests)
head(states)
```

Here are some functions that you may need to use in this lab:

- `nchar()`
- `tolower()`
- `toupper()`
- `casefold()`
- `paste()`
- `paste0()`
- `substr()`


## Number of characters

`nchar()` allows you to count the number of characters in a string. Use it 
on `states` to get the number of characters of each state:

```{r nchar}
# number of characters
nchar(states)
```

### Case folding

There are 3 functions to do case-folding: `tolower()`, `toupper()`, and `casefold()`. Apply each function on `states` to see what happens:

```{r tolower}
# to lower case
tolower(states)
```

```{r toupper}
# to upper case
toupper(states)
```

```{r casefold}
# case folding (upper = TRUE)
casefold(states)
```

```{r}
# case folding (upper = FALSE)
casefold(states)
```


### Length of State Names

We just saw how to use `nchar()` to count the number of characters in each 
state name:

```{r num-chars}
# number of charcaters
num_chars <- nchar(states)
```

Use the vector `num_chars` to obtain a frequency table called `char_freqs`, 
and then plot the frequencies with a bar chart:

```{r freq-table}
# frequency table
char_freqs <- table(num_chars)
# barchart of number-of-characters
barplot(char_freqs)
```


### Pasting strings

R provides the `paste()` function. This function allows you to __paste__ 
(i.e. append, concatenate) character vectors separated by a blank space:

```{r pumpkin}
paste('Pumpkin', 'Pie')
```

You can give it any number of vector inputs

```{r abcde}
paste('a', 'b', 'c', 'd', 'e')
```

You can change the separator with `sep`

```{r abcde-sep}
paste('a', 'b', 'c', 'd', 'e', sep = '-')
```

`paste()` is vectorized:

```{r vectorized}
paste('a', 1:5, sep = '.')
```

There's a special wrapper around `paste()` called `paste0()` which is 
equivalent to `paste(..., sep = "")`

```{r paste0}
# paste0() -vs- paste(..., sep = "")
paste0('Pumpkin', 'Pie')
paste('Pumpkin', 'Pie', sep = '')

# paste0() is also vectorized
paste0('a', 1:5)
```


__Your Turn:__ Use `paste()` to form a new vector with the first five states 
and their number of characters like this:

`"Alabama = 7"  "Alaska = 6"  "Arizona = 7"  "Arkansas = 8"  "California = 10"`

```{r frist5-states}
# paste names with their num-of-chars
paste(head(states), head(num_chars), sep = " = ")
```

Now use `paste()`'s argument `collapse = ''` to _collapse_ the first five 
states like this:

`"AlabamaAlaskaArizonaArkansasCalifornia"`

```{r collapse}
# collapse first 5 states
paste(head(states), collapse = '')
```


### Substrings

R provieds the function `substr()` to extract substrings in a character vector:

```{r substr-berkeley}
# extract first 3 characters
substr('Berkeley', 1, 3)
```

Use `substr()` to shorten the state names using the first 3-letters:

```{r substr-ex1}
# shorten state names with first 3 characters
substr(states, 1, 3)
```


### Challenge

We already obtained a frequency table `char_freqs` with the counts of state 
names by number of characters. You can use those frequencies to get those 
state-names with 4-characters or 10-characters:

```{r}
# 4-char states
states[num_chars == 4]
```

```{r}
# 10-char states
states[num_chars == 10]
```

You can use `paste()` to join the 4-character states in one single string 
(i.e. _collapsing_) like this---separated by a comma and space---:

```{r}
# collapse 4-char states
paste(states[num_chars == 4], collapse = ", ")
```

__Here's one challenge for you:__ write code (using a for-loop) to obtain a 
list `states_list` containing the collapsed names by number of characters. 
If the number of characters is an even number, then the state names should be 
in capital letters. Otherwise, they should be in lower case letters.

Each list element of `states_list` must be named with the number of characters, 
followed by a dash, followed by the word `chars`: e.g. `'4-chars'`, `'5-chars'`,
etc. In total, `states_list` should have the same length as `char_freqs`.

Here's what `states_list` should look like for the first three elements:

```
$`4-chars`
[1] "IOWA, OHIO, UTAH"

$`5-chars`
[1] "idaho, maine, texas"

$`6-chars`
[1] "ALASKA, HAWAII, KANSAS, NEVADA, OREGON"
```

-----

## Converting from Fahrenheit Degrees

Here are four functions that convert from Fahrenheit degrees to other temperature scales:

```{r}
to_celsius <- function(x = 1) {
  (x - 32) * (5/9)
}

to_kelvin <- function(x = 1) {
  (x + 459.67) * (5/9)
}

to_reaumur <- function(x = 1) {
  (x - 32) * (4/9)
}

to_rankine <- function(x = 1) {
  x + 459.67
}
```

We can use the previous functions to create a more general function `temp_convert()`:

```{r}
temp_convert <- function(x = 1, to = "celsius") {
  switch(to,
         "celsius" = to_celsius(x),
         "kelvin" = to_kelvin(x),
         "reaumur" = to_reaumur(x),
         "rankine" = to_rankine(x))
}

temp_convert(30, 'celsius')
```

`temp_convert()` works fine when the argument `to = 'celsius'`. But what happens 
if you try `temp_convert(30, 'Celsius')` or `temp_convert(30, 'CELSIUS')`?

__Your turn__. Rewrite `temp_convert()` such that the argument `to` can be given 
in upper or lower case letters. For instance, the following three calls should 
be equivalent:

```{r, eval=FALSE}
temp_convert(30, 'celsius')
temp_convert(30, 'Celsius')
temp_convert(30, 'CELSIUS')
```

```{r}
temp_convert <- function(x = 1, to = "celsius") {
  switch(tolower(to),
         "celsius" = to_celsius(x),
         "kelvin" = to_kelvin(x),
         "reaumur" = to_reaumur(x),
         "rankine" = to_rankine(x))
}

```

-----



## Valid Color Names

The function `colors()` returns a vector with the names (in English) of 657 
colors available in R. Write a function `is_color()` to test if a given 
name---in English---is a valid R color. If the provided name is a valid R color, 
`is_color()` returns `TRUE`.  If the provided name is not a valid R color 
`is_color()` returns `FALSE`.

```{r, eval=FALSE}
# your is_color() function
is_color <- function(color = "red") {
  tolower(color) %in% colors()
}



# test it:
is_color('yellow')  # TRUE

is_color('blu')     # FALSE

is_color('turkuoise') # FALSE
```


