---
title: "HW02 - Shot Charts"
author: "Kanak Garg"
date: "3/5/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/')
```

```{r}
library(dplyr)
library(ggplot2)
shot_data <- read.csv(file = "../data/shot-data.csv", stringsAsFactors = FALSE)
```


#Total Shots By Player
```{r}
dat1 <- 
shot_data %>%
  group_by(name) %>%
  select(name) %>%
  summarise(total = length(name)) %>%
  arrange(desc(total))
dat1
```


```{r}
dat2 <- 
shot_data %>%
  group_by(name) %>%
  select(name, shot_type, shot_made_flag) %>%
  summarise(
    total = length(name),
    made = length(shot_made_flag[shot_made_flag == "made shot"]),
    perc_made = made/total) %>%
  arrange(desc(total))
dat2
```
```{r}
dat3 <- 
shot_data %>%
  group_by(name) %>%
  select(name, shot_type, shot_made_flag) %>%
  summarise(
    total2pt = length(shot_type[shot_type=="2PT Field Goal"]),
    made2pt = length(shot_made_flag[shot_made_flag == "made shot" & shot_type == "2PT Field Goal"]),
    perc2pt = made2pt/total2pt) %>%
  arrange(desc(total2pt))
dat3
```

```{r}
dat4 <- 
shot_data %>%
  group_by(name) %>%
  select(name, shot_type, shot_made_flag) %>%
  summarise(
    total3pt = length(shot_type[shot_type=="3PT Field Goal"]),
    made3pt = length(shot_made_flag[shot_made_flag == "made shot" & shot_type == "3PT Field Goal"]),
    perc3pt = made3pt/total3pt) %>%
  arrange(desc(total3pt))
dat4
```

```{r}
dat5 <- 
shot_data %>%
  group_by(shot_distance) %>%
  select(shot_distance, shot_made_flag) %>%
  summarise(
    made_shot_prop = length(shot_made_flag[shot_made_flag == "made shot"])/n()) %>%
  arrange(shot_distance)
dat5
```

```{r}
ggplot(data = dat5) + geom_point(aes(x = shot_distance, y = made_shot_prop)) +
  ggtitle('Shot by Distance') 
```

* What do you observe?  
There is a definite correlation between shot distance and the the proportion of shots made. As the distance increases, shot proportion follows a downward trend. There is an outlier at 51 shot distance.

* Can you confirm that the shorter the distance, the more effective the shots?  
No, as distance decreases, the shot proportion made does not necessarily increase. There are instances where shot proportion decreases as distnace decrases.

* Can you guesstimate a distance threshold beyond which the chance of making a
successful shot is basically null?  
After 34, the shot chances are basically null.

* What distances tend to have a percentage (of making a shot) of 50% or more?  
Distances around 0, 1, 2, 12, 16, 18, 21, 30 and 51 tend to have a shot percentage of 50% or more.

```{r}
dat6 <- 
shot_data %>%
  group_by(name, minute, period) %>%
  select(name, minute, period) %>%
  summarise(
    shots = length(minute))
dat6

ggplot(data = dat6) + aes(x = minute, y = shots, xmin = (period%%2)*24, xmax = (period%%2)*24 +12 , ymin = 0, ymax = max(shots)) + geom_rect(fill = "grey95") + geom_point( col = rgb(73,140,251, maxColorValue = 255) ) +
  scale_x_continuous(breaks = c(12,24,36,48)) + geom_path(col = rgb(73,140,251, maxColorValue = 255)) + 
  labs(y = "total number of shots", title = 'Total number of shots (by minute of occurrence)') +theme_minimal() + facet_wrap(~ name)

```

